Non-parametric approaches to optimal policy are more robust

### 2013-04-01 inst/doc/nonparametric-bayes.Rmd
build drafts, update html support

Though machine learning approaches have begun to appear in the
ecological and conservation literature (Species distribution models),
including the Gaussian process based approach used here [@Munch2005],
they remain unfamiliar and untrusted approaches for most ecologists.
One potential barrier to their adoption is the absence of a framework
for applying machine learning approaches to resource management problems.
Traditional approaches to optimal control (Pontryagin's principle, stochastic
dynamic programming) rely on knowledge of the state equation, usually described
by a simple parametric model. Here we illustrate how a stochastic dynamic
programming algorithm can alternatively be driven by the predictions from
a Gaussian process -- a machine learning approximation to the state dynamics.

<!-- Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
obvious to everyon that it's trivial? -->

<!-- paraphrase advantages of machine learning -->

@@

Given parameters for the function $f$ and probability distribution $Z$, 
along with a given economic model determining the  price/profit $\Pi(X_t, h_t)$
realized in a given year given a choice of harvest $h_t$ and observed stock $X_t$.
This problem can be solved exactly for discretized values of stock $X$ and policy $h$
using stochastic dynamic programming (SDP) [@Mangel1982].  
Problems of this sort underpin much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of $f$,
along with its parameters.  The standard approach [@refs] uses one of a 
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any 
available existing data. Though modifications of this approach mentioned in the 
introduction can permit additional sources of uncertainty such as measurement 
error in the stock assessment, implementation errors in the harvest policy, [@Clark1986, @Roughgarden1996, @Sethi2005]
uncertainty in parameters [@Mangel1985, @Schaupaugh2013] or model structure [@Williams2001, @Athanassoglou2012],
we focus on the simplest example to illustrate the comparison, where only a 
single form of uncertainty (the growth shocks $Z_t$) is present in the underlying
model, making additional modes unnecessary. 
<!--Add comment / apology on single species dynamics -->

@@

The use of Gaussian process (GP) regression (known as Kreging in the geospatial
literature) to formulate a predictive model is relatively new in the
context of modeling dynamical systems [@Kocijan2005] and introduced
in the ecological modeling and fisheries management by @Munch2005.
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

The essense of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<!-- Figure 1 include curves drawn from the posterior density?  -->

<!-- Do we need more specifics on Gaussian process as an approximation
to parametric models? Discussion of Gaussian process vs other machine
learning / forecasting approaches that have less clear statistical
foundations?  If so, does this all belong in the discussion? -->

@@

<!--Should we include any emprical examples? -->
Using data simulated from a specified model rather than empirical data
permits the comparison against the true underlying dynamics, setting 
a bar for the optimal performance possible.

@@

<!-- wow, run-on tangent? -->
Predictive accuracy is not the goal of all modeling, as ecologists
have been observing for as long as they made models (perhaps none
more memorably than @Levins1969).  Mechanistic modeling is at its
most powerful not when it is used to directly forecast future states
but when it provides an understanding of how to approach a problem.
SIR-type models from the epidemiological literature are a good example.
While the simplest SIR models have little predictive power over the
outbreak intensity or timing at a particular location, they provide a
powerful understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate every
member of the population to stop the spread, but rather it suffices to
vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.

@@

<!-- Jargony and vague -->
While we have highlighted certain generic of this problem that allow the
nonparametric approach to excel -- short timescales between successive
observations and actions, the accuracy in the appropriate region of state
space, the ability to express uncertainty outside the observed data --
there are equally several aspects for which the nonparametric approach is
at a relative disadvantage.

<!-- Model complexity -->
In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

<!-- Data complexity. Perhaps too far out of scope... -->
The machine learning approach is also better suited to complex and
disparate data.  Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved.  Only thanks to tandem advances in increasing
computational power and hierarchical statistical methodology have we
been able to tackle such intuitively important complexity (and the
potentially new available data that accompanies it) such as spatial
distribution, heterogeneities of space, time, and individuals, to shift
to ecosystem-based approaches from single-species based approaches.
Without the need to formulate mechanisms, many modern machine learning
algorithms can leverage potential information from all available sources
of data directly.  The algorithms can recognize unanticipated or subtle
patterns in large data sets that enable more accurate predictions than
mechanistic models that are formulated at a more macroscopic level.

<!-- 
* Discuss constant escapement in model, in policies.

@@

* Discussion of alternative related approaches: POMDP/MOMDP
-->

>>>>

<span style="color:#007FFF;"> ***This part is overstated and unnecessary***
Though machine learning approaches have begun to appear in the
ecological and conservation literature (e.g. in species distribution models),
including the Gaussian process based approach used here [@Munch2005],
overall machine learning remains an unfamiliar and somewhat untrusted approach for many ecologists.
One potential barrier to their adoption is the absence of a framework
for applying machine learning approaches to resource management problems. </span>

<span style="color:#007FFF;">***transition sentences!?***</span>

Traditional approaches to optimal control (Pontryagin's principle, stochastic
dynamic programming) rely on knowledge of the state equation, usually described
by a simple parametric model. Here we illustrate how a stochastic dynamic
programming algorithm can alternatively be driven by the predictions from
a Gaussian process -- a machine learning approximation to the state dynamics.

<span style="color:#007FFF;">***Okay, perhaps that's novel, but it's pretty trivial.  Isn't it
obvious to everyon that it's trivial?***</span>

<span style="color:#007FFF;">***paraphrase advantages of machine learning***</span>

@@

Given parameters for the function $f$ and probability distribution $Z$, 
along with a given economic model determining the  price/profit $\Pi(X_t, h_t)$
realized in a given year given a choice of harvest $h_t$ and observed stock $X_t$.
This problem can be solved exactly for discretized values of stock $X$ and policy $h$
using stochastic dynamic programming (SDP) [@Mangel1985]. Problems of this sort underpin much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of $f$,
along with its parameters.  The standard approach [@refs] uses one of a 
handful of common parametric models representing the stock-recruitment
relationship, usually after estimating the model parameters from any 
available existing data. Though modifications of this approach mentioned in the 
introduction can permit additional sources of uncertainty such as measurement 
error in the stock assessment, implementation errors in the harvest policy, [@Clark1986; @Roughgarden1996; @Sethi2005]
uncertainty in parameters [@Mangel1985; @Schapaugh2013] or model structure [@Williams2001; @Athanassoglou2012],
we focus on the simplest example to illustrate the comparison, where only a 
single form of uncertainty (the growth shocks $Z_t$) is present in the underlying
model, making additional modes unnecessary. 
<span style="color:#007FFF;">***Add comment / apology on single species dynamics?*** </span>

@@

The use of Gaussian process (GP) regression (or "kreging" in the geospatial
literature) to formulate a predictive model is relatively new in the
context of modeling dynamical systems [@Kocijan2005] and introduced
in the ecological modeling and fisheries management by @Munch2005.
An accessible and thorough introduction to the formulation and use of
GPs can be found in @Rasmussen2006.

The essense of the GP approach can be captured in the
following thought experiment: An exhaustive parametric approach to the
challenge of structural uncertainty might proceed by writing down all
possible functional forms for the underlying dynamical system with all
possible parameter values for each form, and then consider searching
over this huge space to select the most likely model and parameters;
or using a Bayesian approach, assign priors to each of these possible
models and infer the posterior distribution of possible models. The
GP approach can be thought of as a computationally
efficient approximation to this approach. GPs represent
a large class of models that can be though of as capturing or reasonably
approximating the set of models in this collection.  By modeling at the
level of the process, rather than the level of parametric equation,
we can more concisely capture the possible behavior of these curves.
In place of a parametric model of the dynamical system, the GP 
approach postulates a prior distribution of (n-dimensional)
curves that can be though of as approximations to a range of possible
(parametric) models that might describe the data. The GP allows us
to consider a set of possible curves simultaneously.  
<span style="color:#007FFF;"> ***Figure 1 include curves drawn from the posterior density?*** </span>

<span style="color:#007FFF;">***Do we need more specifics on Gaussian process as an approximation
to parametric models? Discussion of Gaussian process vs other machine
learning / forecasting approaches that have less clear statistical
foundations?  If so, does this all belong in the discussion?***</span>

@@

<span style="color:#007FFF;">***Should we include any emprical examples?***</span>
Using data simulated from a specified model rather than empirical data
permits the comparison against the true underlying dynamics, setting 
a bar for the optimal performance possible.

@@

<span style="color:#007FFF;"> wow, run-on tangent? </span>
Predictive accuracy is not the goal of all modeling, as ecologists
have been observing for as long as they made models (perhaps none
more memorably than @Levins1969).  Mechanistic modeling is at its
most powerful not when it is used to directly forecast future states
but when it provides an understanding of how to approach a problem.
SIR-type models from the epidemiological literature are a good example.
While the simplest SIR models have little predictive power over the
outbreak intensity or timing at a particular location, they provide a
powerful understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate every
member of the population to stop the spread, but rather it suffices to
vaccinate a sufficient fraction of the population to reduce $R_0$ below 1.

@@

<span style="color:#007FFF;"> Jargony and vague </span>
While we have highlighted certain generic of this problem that allow the
nonparametric approach to excel -- short timescales between successive
observations and actions, the accuracy in the appropriate region of state
space, the ability to express uncertainty outside the observed data --
there are equally several aspects for which the nonparametric approach is
at a relative disadvantage.

<span style="color:#007FFF;"> Model complexity </span>
In this simulated example, the underlying
dynamics are truly governed by a simple parametric model, allowing
the parametric approaches to be more accurate.  Similarly, because the
dynamics are  one-dimensional dynamics and lead to  stable nodes (rather
than other attractors such as limit-cycles resulting in oscillations),
the training data provides relatively limited information about the
dynamics.  For these reasons, we anticipate that in higher-dimensional
examples characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

<span style="color:#007FFF;"> Data complexity. Perhaps too far out of scope... </span>
The machine learning approach is also better suited to complex and
disparate data.  Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved.  Only thanks to tandem advances in increasing
computational power and hierarchical statistical methodology have we
been able to tackle such intuitively important complexity (and the
potentially new available data that accompanies it) such as spatial
distribution, heterogeneities of space, time, and individuals, to shift
to ecosystem-based approaches from single-species based approaches.
Without the need to formulate mechanisms, many modern machine learning
algorithms can leverage potential information from all available sources
of data directly.  The algorithms can recognize unanticipated or subtle
patterns in large data sets that enable more accurate predictions than
mechanistic models that are formulated at a more macroscopic level.

<span style="color:#007FFF;"> 
* Discuss constant escapement in model, in policies.

@@

* Discussion of alternative related approaches: POMDP/MOMDP
</span>

### 2013-04-01 inst/doc/nonparametric-bayes.tex
build drafts, update html support

The ecological dynamics of most systems of management interest are
typically both complex and unknown {[}@refs{]}. Despite this,
quantitative policy optimization in these systems has almost exclusively
been based on simple, mechanistically motivated parametric models. There
are several reasons for this. First, limited available data almost
always percludes any accurate estimate of more complicated models.
Second, the computational demands of determining an optimal stategy in a
sequential decision-theory problem suffer from the ``curse of
dimensionality'' that makes\\it difficult to consider more complicated
models, states, or actions. Unfortunately, these challenges also make it
particularly difficult to reflect the true uncertainty either in the
parameters of the given models (parametric uncertainty) or in the
structure of the models themselves (structural uncertainty).\\Further
unknowns such as measurement uncertainty, parameter uncertainty,
unobserved states, knowledge of boundary conditions, etc. further
compound the issue. Though a hierarchical Bayesian approach provides a
natural way to address these from a statistical standpoint, formulating
reasonable parametric descriptions of each form of uncertainty is a
challenging task in itself, let alone the computational difficulties of
solving such a system. Cressie et al. (2009) provides a good account of
the successess and challenges of the approach. Applying these approaches
in the management context of sequential decision making, in which
forecasts must be obtained over a range of possible actions and updated
regularly as new information arrives makes such an approach less
feasible still.

@@

Though machine learning approaches have begun to appear in the
ecological and conservation literature (Species distribution models),
including the Gaussian process based approach used here (Munch et al.
2005), they remain unfamiliar and untrusted approaches for most
ecologists. One potential barrier to their adoption is the absence of a
framework for applying machine learning approaches to resource
management problems. Traditional approaches to optimal control
(Pontryagin's principle, stochastic dynamic programming) rely on
knowledge of the state equation, usually described by a simple
parametric model. Here we illustrate how a stochastic dynamic
programming algorithm can alternatively be driven by the predictions
from a Gaussian process -- a machine learning approximation to the state
dynamics.

@@

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) {[}@Mangel1982{]}.\\Problems of this sort underpin
much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach {[}@refs{]} uses
one of a handful of common parametric models representing the
stock-recruitment relationship, usually after estimating the model
parameters from any available existing data. Though modifications of
this approach mentioned in the introduction can permit additional
sources of uncertainty such as measurement error in the stock
assessment, implementation errors in the harvest policy, (Clark and
Kirkwood 1986, , ) uncertainty in parameters (Mangel and Clark 1985,
@Schaupaugh2013) or model structure (B. K. Williams 2001, ), we focus on
the simplest example to illustrate the comparison, where only a single
form of uncertainty (the growth shocks $Z_t$) is present in the
underlying model, making additional modes unnecessary.

@@

The use of Gaussian process (GP) regression (known as Kreging in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in C. E. Rasmussen and Williams
(2006).

The essense of the GP approach can be captured in the following thought
experiment: An exhaustive parametric approach to the challenge of
structural uncertainty might proceed by writing down all possible
functional forms for the underlying dynamical system with all possible
parameter values for each form, and then consider searching over this
huge space to select the most likely model and parameters; or using a
Bayesian approach, assign priors to each of these possible models and
infer the posterior distribution of possible models. The GP approach can
be thought of as a computationally efficient approximation to this
approach. GPs represent a large class of models that can be though of as
capturing or reasonably approximating the set of models in this
collection. By modeling at the level of the process, rather than the
level of parametric equation, we can more concisely capture the possible
behavior of these curves. In place of a parametric model of the
dynamical system, the GP approach postulates a prior distribution of
(n-dimensional) curves that can be though of as approximations to a
range of possible (parametric) models that might describe the data. The
GP allows us to consider a set of possible curves simultaneously.\\

@@

Using data simulated from a specified model rather than empirical data
permits the comparison against the true underlying dynamics, setting a
bar for the optimal performance possible.

@@

Predictive accuracy is not the goal of all modeling, as ecologists have
been observing for as long as they made models (perhaps none more
memorably than Levins (1969)). Mechanistic modeling is at its most
powerful not when it is used to directly forecast future states but when
it provides an understanding of how to approach a problem. SIR-type
models from the epidemiological literature are a good example. While the
simplest SIR models have little predictive power over the outbreak
intensity or timing at a particular location, they provide a powerful
understanding of the spread of an infection in terms of a single,
biologically meaningful parameter: $R_0$, the basic reproductive number.
From the model, it becomes clear that management need not vaccinate
every member of the population to stop the spread, but rather it
suffices to vaccinate a sufficient fraction of the population to reduce
$R_0$ below 1.

@@

While we have highlighted certain generic of this problem that allow the
nonparametric approach to excel -- short timescales between successive
observations and actions, the accuracy in the appropriate region of
state space, the ability to express uncertainty outside the observed
data -- there are equally several aspects for which the nonparametric
approach is at a relative disadvantage.

In this simulated example, the underlying dynamics are truly governed by
a simple parametric model, allowing the parametric approaches to be more
accurate. Similarly, because the dynamics are one-dimensional dynamics
and lead to stable nodes (rather than other attractors such as
limit-cycles resulting in oscillations), the training data provides
relatively limited information about the dynamics. For these reasons, we
anticipate that in higher-dimensional examples characteristic of
ecosystem management problems that the machine learning approach will
prove even more valuable.

The machine learning approach is also better suited to complex and
disparate data. Incorporating various sources of information into
mechanistic models can be an immensely difficult due to the increased
complexity involved. Only thanks to tandem advances in increasing
computational power and hierarchical statistical methodology have we
been able to tackle such intuitively important complexity (and the
potentially new available data that accompanies it) such as spatial
distribution, heterogeneities of space, time, and individuals, to shift
to ecosystem-based approaches from single-species based approaches.
Without the need to formulate mechanisms, many modern machine learning
algorithms can leverage potential information from all available sources
of data directly. The algorithms can recognize unanticipated or subtle
patterns in large data sets that enable more accurate predictions than
mechanistic models that are formulated at a more macroscopic level.

@@

>>>>

The ecological dynamics of most systems of management interest are
typically both complex and unknown {[}@refs{]}. Despite this,
quantitative policy optimization in these systems has almost exclusively
been based on simple, mechanistically motivated parametric models. There
are several reasons for this. First, limited available data almost
always percludes any accurate estimate of more complicated models.
Second, the computational demands of determining an optimal stategy in a
sequential decision-theory problem suffer from the ``curse of
dimensionality'' that makes it difficult to consider more complicated
models, states, or actions. Unfortunately, these challenges also make it
particularly difficult to reflect the true uncertainty either in the
parameters of the given models (parametric uncertainty) or in the
structure of the models themselves (structural uncertainty). Further
unknowns such as measurement uncertainty, parameter uncertainty,
unobserved states, knowledge of boundary conditions, etc. further
compound the issue. Though a hierarchical Bayesian approach provides a
natural way to address these from a statistical standpoint, formulating
reasonable parametric descriptions of each form of uncertainty is a
challenging task in itself, let alone the computational difficulties of
solving such a system. Cressie et al. (2009) provides a good account of
the successess and challenges of the approach. Applying these approaches
in the management context of sequential decision making, in which
forecasts must be obtained over a range of possible actions and updated
regularly as new information arrives makes such an approach less
feasible still.

@@

 \textbf{This part is overstated and unnecessary} Though machine
learning approaches have begun to appear in the ecological and
conservation literature (e.g.~in species distribution models), including
the Gaussian process based approach used here (Munch et al. 2005),
overall machine learning remains an unfamiliar and somewhat untrusted
approach for many ecologists. One potential barrier to their adoption is
the absence of a framework for applying machine learning approaches to
resource management problems.

 transition sentences!?

Traditional approaches to optimal control (Pontryagin's principle,
stochastic dynamic programming) rely on knowledge of the state equation,
usually described by a simple parametric model. Here we illustrate how a
stochastic dynamic programming algorithm can alternatively be driven by
the predictions from a Gaussian process -- a machine learning
approximation to the state dynamics.

 Okay, perhaps that's novel, but it's pretty trivial. Isn't it obvious
to everyon that it's trivial?

 paraphrase advantages of machine learning

@@

Given parameters for the function $f$ and probability distribution $Z$,
along with a given economic model determining the price/profit
$\Pi(X_t, h_t)$ realized in a given year given a choice of harvest $h_t$
and observed stock $X_t$. This problem can be solved exactly for
discretized values of stock $X$ and policy $h$ using stochastic dynamic
programming (SDP) (Mangel and Clark 1985). Problems of this sort
underpin much marine fisheries management today.

A crux of this approach is correctly specifying the functional form of
$f$, along with its parameters. The standard approach {[}@refs{]} uses
one of a handful of common parametric models representing the
stock-recruitment relationship, usually after estimating the model
parameters from any available existing data. Though modifications of
this approach mentioned in the introduction can permit additional
sources of uncertainty such as measurement error in the stock
assessment, implementation errors in the harvest policy, (Clark and
Kirkwood 1986; Roughgarden and Smith 1996; Sethi et al. 2005)
uncertainty in parameters (Mangel and Clark 1985; Schapaugh and Tyre
2013) or model structure (B. K. Williams 2001; Athanassoglou and
Xepapadeas 2012), we focus on the simplest example to illustrate the
comparison, where only a single form of uncertainty (the growth shocks
$Z_t$) is present in the underlying model, making additional modes
unnecessary. Add comment / apology on single species dynamics

@@

The use of Gaussian process (GP) regression (or ``kreging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005) and
introduced in the ecological modeling and fisheries management by Munch
et al. (2005). An accessible and thorough introduction to the
formulation and use of GPs can be found in C. E. Rasmussen and Williams
(2006).

The essense of the GP approach can be captured in the following thought
experiment: An exhaustive parametric approach to the challenge of
structural uncertainty might proceed by writing down all possible
functional forms for the underlying dynamical system with all possible
parameter values for each form, and then consider searching over this
huge space to select the most likely model and parameters; or using a
Bayesian approach, assign priors to each of these possible models and
infer the posterior distribution of possible models. The GP approach can
be thought of as a computationally efficient approximation to this
approach. GPs represent a large class of models that can be though of as
capturing or reasonably approximating the set of models in this
collection. By modeling at the level of the process, rather than the
level of parametric equation, we can more concisely capture the possible
behavior of these curves. In place of a parametric model of the
dynamical system, the GP approach postulates a prior distribution of
(n-dimensional) curves that can be though of as approximations to a
range of possible (parametric) models that might describe the data. The
GP allows us to consider a set of possible curves simultaneously.\\
Figure 1 include curves drawn from the posterior density?

 Do we need more specifics on Gaussian process as an approximation to
parametric models? Discussion of Gaussian process vs other machine
learning / forecasting approaches that have less clear statistical
foundations? If so, does this all belong in the discussion?

@@

Should we include any emprical examples? Using data simulated from a
specified model rather than empirical data permits the comparison
against the true underlying dynamics, setting a bar for the optimal
performance possible.

@@

 wow, run-on tangent? Predictive accuracy is not the goal of all
modeling, as ecologists have been observing for as long as they made
models (perhaps none more memorably than Levins (1969)). Mechanistic
modeling is at its most powerful not when it is used to directly
forecast future states but when it provides an understanding of how to
approach a problem. SIR-type models from the epidemiological literature
are a good example. While the simplest SIR models have little predictive
power over the outbreak intensity or timing at a particular location,
they provide a powerful understanding of the spread of an infection in
terms of a single, biologically meaningful parameter: $R_0$, the basic
reproductive number. From the model, it becomes clear that management
need not vaccinate every member of the population to stop the spread,
but rather it suffices to vaccinate a sufficient fraction of the
population to reduce $R_0$ below 1.

@@

 Jargony and vague While we have highlighted certain generic of this
problem that allow the nonparametric approach to excel -- short
timescales between successive observations and actions, the accuracy in
the appropriate region of state space, the ability to express
uncertainty outside the observed data -- there are equally several
aspects for which the nonparametric approach is at a relative
disadvantage.

 Model complexity In this simulated example, the underlying dynamics are
truly governed by a simple parametric model, allowing the parametric
approaches to be more accurate. Similarly, because the dynamics are
one-dimensional dynamics and lead to stable nodes (rather than other
attractors such as limit-cycles resulting in oscillations), the training
data provides relatively limited information about the dynamics. For
these reasons, we anticipate that in higher-dimensional examples
characteristic of ecosystem management problems that the machine
learning approach will prove even more valuable.

 Data complexity. Perhaps too far out of scope\ldots{} The machine
learning approach is also better suited to complex and disparate data.
Incorporating various sources of information into mechanistic models can
be an immensely difficult due to the increased complexity involved. Only
thanks to tandem advances in increasing computational power and
hierarchical statistical methodology have we been able to tackle such
intuitively important complexity (and the potentially new available data
that accompanies it) such as spatial distribution, heterogeneities of
space, time, and individuals, to shift to ecosystem-based approaches
from single-species based approaches. Without the need to formulate
mechanisms, many modern machine learning algorithms can leverage
potential information from all available sources of data directly. The
algorithms can recognize unanticipated or subtle patterns in large data
sets that enable more accurate predictions than mechanistic models that
are formulated at a more macroscopic level.

 * Discuss constant escapement in model, in policies.

\begin{itemize}
\item
  Limitations of this comparison: Are the maximum-likelihood solutions a
  straw man?
\item
  Discussion of alternative related approaches: POMDP/MOMDP 
\end{itemize}

@@

Schapaugh, Adam W., and Andrew J. Tyre. 2013. ``Accounting for
parametric uncertainty in Markov decision processes.'' \emph{Ecological
Modelling} 254 (apr): 15--21. doi:10.1016/j.ecolmodel.2013.01.003.
\url{http://linkinghub.elsevier.com/retrieve/pii/S0304380013000306}.
